Explaining the reportedly over-massive black holes in early-type galaxies with intermediate-scale discs

### 2015-10-30 papers/ellicular/ellicular.tex
fanculo

\usepackage{longtable}
\usepackage[]{graphicx}
\usepackage{amsmath}
\usepackage{natbib}
\bibliographystyle{mnras}

@@

Building on the investigation of \cite{rixwhite1990}, 

@@

This sequence is analogue to that illustrated in Figure 2 of \cite{cappellari2011}, 

@@

In general, stellar discs are intrinsically flat and close to circular; 

@@

\subsection{ES galaxies in the Hubble grid}
{\bf The existence of intermediate-scale discs has led to the suggestion of a likely continuum of disc sizes in early-type galaxies 
\citep{simiendevaucouleurs1986,capaccioli1988,bender1989,simienmichard1990,scorza1993,scorzavandenBosch1998} 
as opposed to a dichotomy of say nuclear discs versus large-scale discs.  
The presence of these intermediate-scale discs blurs the distinction between elliptical (E) and lenticular (S0) galaxies, 
or central slow-rotators (SR) versus central fast-rotators (FR), see \citeauthor{cappellari2011} (\citeyear{cappellari2011}, their Figure 1 and 2).   
There is a need for more than two bins (i.e.~nuclear vs.~large-scale disc, E vs.~S0, SR vs.~FR), 
and a need for a continuum in the (Hubble-Jeans)-like classification 
schemes\footnote{As noted by \cite{vandenbergh1997} and \cite{sandage2005}, 
it was Sir James Jeans (1928) who introduced the (tuning fork)-shaped diagram that encapsulated \citeauthor{Hubble1926}'s (\citeyear{Hubble1926}) 
elliptical-spiral sequence -- itself motivated by the earlier work of \cite{Jeans1919}.}.

It has been argued that the classification scheme for "elliptical" galaxies should not be 
their apparent axis ratio as seen on the plane of the sky (originally E0--E7, then E0--E4,  see Hubble 1936; Sandage 1961; 
Liller 1966; Cappellari et al.\ 2011), because this depends on the viewing angle rather than being intrinsic to the galaxy.  
Obviously when dealing with intermediate-scale discs, 
the boxy/discy shape of the isophotes is also a function of both radius and disc inclination, 
making single isophotal shape parameters problematic and particularly inappropriate for the ES class.  
Section 4 of Kormendy & Bender (1996) points out additional complications in regard to this.  
Instead, early-type galaxies would be better quantified by their spheroid-to-total flux ratio, 
with a continuum from pure elliptical galaxies to disc-dominated lenticular galaxies (e.g., Capaccioli et al.\ 1988, and references therein).   
This spheroid-to-total flux (or better mass) ratio is widely recognised as an important quantity, 
and its broad range observed in the lenticular galaxies led Cappellari et al.\ (2011b) to present the "Hubble comb".   
Building on van den Bergh's (1976) diagram in which both late-type galaxies (i.e., spiral galaxies) 
and early-type galaxies form separate prongs of an expanded Hubble-Jeans tuning fork, 
with an additional prong for disc galaxies hosting anaemic spiral patterns, 
Cappellari et al. (2011b, see also Kormendy & Bender 2012) presented a scheme in which the morphological type (a, b, c, etc.) 
changed if the spheroid-to-total ratio changed.  
However the primary criteria used to assign the spiral galaxy type over the last half century (Sandage 1961) 
has been the nature of the spiral arms --- the extent to which they are unwound, pitch angle, 
and the degree of resolution --- rather than the spheroid-to-total ratio.   
Indeed, in the Hubble Atlas of Galaxies, Sandage (1961) notes that Sa type galaxies exist with both small and large bulges.

The grid seen in Figure~6 and discussed in Graham (2014) was used by Freeman (1970, his Figure~9), 
Boroson (1981), Kent (1985), Kodaira et al. (1986), Simien & de Vaucouleurs (1986) and others, 
and allows for the varying spheroid-to-total ratio among each galaxy type.   
Rather than using the extended classification scheme for early-type galaxies (E - E$^+$ - L$^-$ - L$^o$ - L$^+$-S0, 
corresponding to the numerical T-type ranging from -5 to 0, e.g., de Vaucouleurs et al.\ 1964, Heidmann et al.\ 1972), 
we have simply grouped them all together.   While, collectively, spiral galaxies typically display a range of spheroid-to-total flux 
ratios from 0 to 3/4 (e.g. Graham & Worley 2008, and references therein), 
lenticular galaxies commonly display a range from 3/4 to 0.1 (Laurikainen et al. 2010), and the ES class have larger ratios.  
The purpose here is not to provide a complete classification scheme, 
but to simply show where the early-type galaxies with intermediate-scale discs (ES) reside relative to the other 
Hubble galaxy types in this familiar diagram which has tended not to include the elliptical galaxies to date.

In Figure 6, can you please move the ES galaxy to a spheroid-to-total ratio of 0.75.  
Common lenticular galaxies with large-scale discs are known to have ratios up to ~0.75.
Moving the ES galaxy in Figure 6 will therefore better match with the idea that these are 
intermediate between S0 and E galaxies. }

\begin{figure}[h]
\begin{center}
\includegraphics[width=\columnwidth]{Hubblegrid.pdf}
\caption{The Hubble grid showing the location of the ES galaxies, intermediate between
elliptical (E) and lenticular (S0) galaxies, relative to the spiral galaxies.  }
\label{fig:grid}
\end{center}
\end{figure}

\section{Summary and conclusions}
{\bf Early-type galaxies display a broad distribution of spheroid-to-total flux ratios (e.g.~\citealt{cappellari2011}), 
going from disc-less, ``pure'' elliptical galaxies (slow rotators) 
to disc-dominated lenticular galaxies (central fast rotators that continue to be fast rotating also beyond one half-light radius). 
%\citep{michard1984,djorgovski1985,carter1987,capaccioli1987,capaccioli1988,franx1989,nieto1988,nieto1991,rixwhite1990,rixwhite1992,
%vandenbergh1990,bender1990,scorzabender1990,scorzabender1995,
%simienmichard1990,cinzanovandermarel1993,donofrio1995,graham1998fornax,scorza1998,scorzavandenbosch1998,
%bendersaglia1999}\footnote{Many past studies were hampered by bulge/disc decompositions which forced an $R^{1/4}$ model for the bulge component 
%and thus obtained spurious bulge-to-disc ratios, and from the \emph{IRAF} task {\tt ellipse} which does not report the correct deviations of isophotes 
%from ellipses \citep{ciambur2015}.}. 
In between these two extremes lie galaxies with intermediate-scale discs 
(spheroid dominated central fast rotators that become slow rotating in their outer regions), 
i.e.~discs of kiloparsec-size that remain ``embedded'' within the spheroidal component of the galaxy 
and do not dominate the galaxy light at large radii as large-scale discs do. }
While this is likely known to some readers, 
the surge of papers presenting galaxy decompositions which are not aware of this reality 
has created a pressing need for this reminder. 
We have shown that the light distribution of galaxies with intermediate-scale discs can be accurately described 
with a simple spheroid + disc (+ optional nuclear component) model, 
without the need for the addition of a bright envelope-component. \\
Our decompositions correctly reproduce both the photometric (surface brightness and ellipticity profiles) 
and kinematic (specific angular momentum profile) properties of nine intermediate-scale disc galaxies. 
Four of these nine galaxies (Mrk 1216, NGC 1271, NGC 1277, NGC 1332) and one additional elliptical galaxy (NGC 4291) 
had previously been claimed to be extreme outliers in the (black hole mass)-(spheroid mass) diagram. 
However, here we have demonstrated that, when correctly modelled, 
these five galaxies all reside well within the scatter of the correlation, 
i.e. they do not host over-massive black holes. 
This serves to strengthen the (black hole mass)-(spheroid mass) relation, 
and rules out the need for exotic formation scenarios.

\section{Acknowledgments}
This research was supported by Australian Research Council funding through grant FT110100263. 
GS is grateful to Matteo Fossati, Luca Cortese and Giuseppe Gavazzi for useful comments and discussion. 
% referee thank you!!
This work is based on observations made with the IRAC instrument \citep{fazio2004IRAC} on-board the Spitzer Space Telescope, 
which is operated by the Jet Propulsion Laboratory, California Institute of Technology under a contract with NASA, 
and also on observations made with the NASA/ESA Hubble Space Telescope, 
and obtained from the Hubble Legacy Archive, 
which is a collaboration between the Space Telescope Science Institute (STScI/NASA), 
the Space Telescope European Coordinating Facility (ST-ECF/ESA) and the Canadian Astronomy Data Centre (CADC/NRC/CSA).
This research has made use of the GOLDMine database \citep{goldmine} and the NASA/IPAC Extragalactic Database (NED) 
which is operated by the Jet Propulsion Laboratory, California Institute of Technology, 
under contract with the National Aeronautics and Space Administration.

>>>>

\usepackage{longtable}
\usepackage[]{graphicx}
\usepackage{amsmath}
\usepackage{natbib}
\usepackage{soul}
\bibliographystyle{mnras}

@@

Building on the investigations in works such as \cite{liller1966}, \cite{Jedrzejewski1987INPROCEEDINGS} and \cite{rixwhite1990}, 

@@

This sequence is analogue to that illustrated in Figure 2 of \cite{cappellari2011kmdr}, 

@@

In general, stellar discs are intrinsically flat and close to circular \citep{Andersen2001,AndersenBershady2002}; 

@@

\subsection{ES galaxies in the Hubble grid}
{\bf The existence of intermediate-scale discs has led to the suggestion of a likely continuum of disc sizes in early-type galaxies 
\citep{simiendevaucouleurs1986,capaccioli1988,bender1989,simienmichard1990,scorza1993,scorzavandenBosch1998} 
as opposed to a dichotomy of say nuclear discs versus large-scale discs.  
The presence of these intermediate-scale discs blurs the distinction between elliptical (E) and lenticular (S0) galaxies, 
or slow-rotators (SR) versus fast-rotators (FR), 
see \citeauthor{cappellari2011kmdr} (\citeyear{cappellari2011kmdr}, their Figure 1 and 2).   
There is a need for more than two bins (i.e.~nuclear vs.~large-scale disc, E vs.~S0, SR vs.~FR), 
and a need for a continuum in the (Hubble-Jeans)-like classification 
schemes\footnote{As noted by \cite{vandenbergh1997} and \cite{sandage2005}, 
it was Sir James \cite{jeans1928} who introduced the (tuning fork)-shaped diagram that encapsulated \citeauthor{Hubble1926}'s (\citeyear{Hubble1926}) 
elliptical-spiral sequence -- itself motivated by the earlier work of \cite{Jeans1919}.}.     \\
It has been argued that the classification scheme for ``elliptical'' galaxies should not be 
their apparent axis ratio as seen on the plane of the sky (originally E0--E7, see \citealt{hubble1936,sandage1961}, 
then E0--E4, see \citealt{liller1966} and also \citealt{gorbachev1970}), 
because this depends on the viewing angle rather than being intrinsic to the galaxy.  
Obviously when dealing with intermediate-scale discs, 
the boxy/discy shape of the isophotes is also a function of both radius and disc inclination, 
making single isophotal shape parameters problematic and particularly inappropriate for the ES class.  
Section 4 of \cite{kormendybender1996} points out additional complications in regard to this.  
Instead, early-type galaxies would be better quantified by their spheroid-to-total flux ratio, 
with a continuum from pure elliptical galaxies to disc-dominated lenticular galaxies (e.g.~\citealt{capaccioli1988}, and references therein).   
This spheroid-to-total flux (or better mass) ratio is widely recognised as an important quantity, 
and its broad range observed in the lenticular galaxies led \cite{cappellari2011kmdr} to present the ``Hubble comb''.   
Building on \citeauthor{vandenbergh1976}'s (\citeyear{vandenbergh1976}) diagram in which both late-type galaxies (i.e.~spiral galaxies) 
and early-type galaxies form separate prongs of an expanded Hubble-Jeans tuning fork, 
with an additional prong for disc galaxies hosting anaemic spiral patterns, 
\citeauthor{cappellari2011kmdr} (\citeyear{cappellari2011kmdr}, see also \citealt{kormendybender2012}) 
presented a scheme in which the morphological type (a, b, c, etc.) 
changed if the spheroid-to-total ratio changed.  
However the primary criteria used to assign the spiral galaxy type over the last half century \citep{sandage1961} 
has been the nature of the spiral arms -- the extent to which they are unwound, pitch angle, 
and the degree of resolution -- rather than the spheroid-to-total ratio.   
Indeed, in the Hubble Atlas of Galaxies, \cite{sandage1961} notes that Sa type galaxies exist with both small and large bulges 
(see also \citealt{VorontsovvelYaminovArkhipova1962} and \citealt{Vorontsovvelyaminovnoskova1973}). \\
The grid seen in Figure \ref{fig:grid} and discussed in \cite{graham2014review} 
was used by \citeauthor{freeman1970} (\citeyear{freeman1970}, his Figure 9), 
\cite{boroson1981}, \cite{kent1985}, \cite{kodaira1986}, \cite{simiendevaucouleurs1986} and others, 
and allows for the varying spheroid-to-total ratio among each galaxy type.   
Rather than using the extended classification scheme for early-type galaxies (E -- E$^+$ -- L$^-$ -- L$^o$ -- L$^+$ -- S0, 
corresponding to the numerical T-type ranging from -5 to 0, e.g.~\citealt{devaucouleurs1964,heidmann1972}), 
we have simply grouped them all together.  
While, collectively, spiral galaxies typically display a range of spheroid-to-total flux 
ratios from $0$ to $\approx 3/4$ (e.g.~\citealt{grahamworley2008}, and references therein), 
lenticular galaxies commonly display a range from $\approx 3/4$ to $\approx 0.1$ \citep{laurikainen2010}, and the ES class have larger ratios.  
The purpose here is not to provide a complete classification scheme, 
but to simply show where the early-type galaxies with intermediate-scale discs (ES) reside relative to the other 
Hubble galaxy types in this familiar diagram which has tended not to include the elliptical galaxies to date. }

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{Hubblegrid.pdf}
\caption{The Hubble grid showing the location of the ES galaxies, intermediate between
elliptical (E) and lenticular (S0) galaxies, relative to the spiral galaxies.  }
\label{fig:grid}
\end{center}
\end{figure}

\section{Summary and conclusions}
{\bf Early-type galaxies display a broad distribution of spheroid-to-total flux ratios (e.g.~\citealt{cappellari2011kmdr}), 
going from disc-less, ``pure'' elliptical galaxies (slow rotators) 
to disc-dominated lenticular galaxies (central fast rotators that continue to be fast rotating also beyond one half-light radius). 
In between these two extremes lie galaxies with intermediate-scale discs 
(spheroid dominated central fast rotators that become slow rotating in their outer regions), 
i.e.~discs of kiloparsec-size that remain ``embedded'' within the spheroidal component of the galaxy 
and do not dominate the galaxy light at large radii as large-scale discs do. }
While this is likely known to some readers, 
the surge of papers presenting galaxy decompositions which are not aware of this reality 
has created a pressing need for this reminder. 
We have shown that the light distribution of galaxies with intermediate-scale discs can be accurately described 
with a simple spheroid + disc (+ optional nuclear component) model, 
without the need for the addition of a bright envelope-component. \\
Our decompositions correctly reproduce both the photometric (surface brightness and ellipticity profiles) 
and kinematic (specific angular momentum profile) properties of nine intermediate-scale disc galaxies. 
Four of these nine galaxies (Mrk 1216, NGC 1271, NGC 1277, NGC 1332) and one additional elliptical galaxy (NGC 4291) 
had previously been claimed to be extreme outliers in the (black hole mass)-(spheroid mass) diagram. 
However, here we have demonstrated that, when correctly modelled, 
these five galaxies all reside well within the scatter of the correlation, 
i.e. they do not host over-massive black holes. 
This serves to strengthen the (black hole mass)-(spheroid mass) relation, 
and rules out the need for exotic formation scenarios.

\section{Acknowledgments}
This research was supported by Australian Research Council funding through grant FT110100263. 
GS is grateful to Matteo Fossati, Luca Cortese and Giuseppe Gavazzi for useful comments and discussion. 
We are grateful to our anynimous referee for their very careful review of our paper, 
and for the comments, corrections and suggestions that ensued. 
This work is based on observations made with the IRAC instrument \citep{fazio2004IRAC} on-board the Spitzer Space Telescope, 
which is operated by the Jet Propulsion Laboratory, California Institute of Technology under a contract with NASA, 
and also on observations made with the NASA/ESA Hubble Space Telescope, 
and obtained from the Hubble Legacy Archive, 
which is a collaboration between the Space Telescope Science Institute (STScI/NASA), 
the Space Telescope European Coordinating Facility (ST-ECF/ESA) and the Canadian Astronomy Data Centre (CADC/NRC/CSA).
This research has made use of the GOLDMine database \citep{goldmine} and the NASA/IPAC Extragalactic Database (NED) 
which is operated by the Jet Propulsion Laboratory, California Institute of Technology, 
under contract with the National Aeronautics and Space Administration.
